---
title:  "DA2: Assignment 1"
author: "Abduvosid Malikov"
date: "29 November"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

## Introduction

This is the Assignment 1 for `Data Analysis 2`  and `Coding` course. The aim of this project is  to analyse the pattern of association between *registered COVID-19 cases* and *registered number of death* due to COVID-19 on 10-31-2020. 
As a source, [covid-19 data](https://github.com/CSSEGISandData/COVID-19/blob/master/csse_covid_19_data/csse_covid_19_daily_reports/10-31-2020.csv) from CSSE^[Center for Systems Science and Engineering] was used. 

``` {r, warning = FALSE, message = FALSE, include = FALSE}
library(knitr)
library(tidyverse)
require(scales)
library(lspline)
library(estimatr)
library(texreg)
library(ggthemes)

my_file <- "../data/clean/covid_pop_09_11_2020_clean.csv"

# make dataframe from a file
df <- read_csv(my_file)
```

``` {r, echo = FALSE, warning = FALSE, message = FALSE}

df %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~key, scales = "free") +
  geom_histogram()+
  theme_wsj() + 
  scale_fill_wsj()


# summary( df )
```

``` {r, warning = FALSE, message = FALSE, include = FALSE}

#' Check and report your distributions for y and x variables: 
#' use histograms and summary statistics table 
#' (mean, median, min, max, standard deviation)

summary_stats_death <- df %>% summarise(
  mean = mean(death),
  median = median(death),
  std = sd(death),
  min = min(death),
  max = max(death))

summary_stats_confirmed <- df %>% summarise(
  mean = mean(confirmed),
  median = median(confirmed),
  std = sd(confirmed),
  min = min(confirmed),
  max = max(confirmed))
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
kable(summary_stats_death, caption = "Summary statistics on number of death")

kable(summary_stats_confirmed, caption = "Summary statistics on number of confirmed cases") 

```


## level - level
does not give better interpretation
There are some influential observations that governing the conditional mean. 

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# No_of_death - No_of_confirmed: level-level model without scaling
# model is not capturing the data very well. model does not fit the data
ggplot( df , aes(x = confirmed, y = death)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "Number of registered case (31 October 2020)",y = "Number of registered death") 

```

## level y - log x
capturing the non-linearity in data
```{r, echo = FALSE, warning = FALSE, message = FALSE}
ggplot( df , aes(x = confirmed, y = death)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "Number of registered case (31 October 2020, ln scale)",y = "Number of registered death") +
  scale_x_continuous( trans = log_trans() )
```

## log y - level x

non-linear

Figure 8.6 A, p214

Most of the observations are very close to each other at the bottom of the distribution of Number of registered cases.



```{r, echo = FALSE, warning = FALSE, message = FALSE}

ggplot( df , aes(x = confirmed, y = death)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "Number of registered case (31 October 2020)",y = "Number of registered death (ln scale)") +
  scale_y_continuous( trans = log_trans() )
```

## log y - log x

Taking log of number of registered death and log of number of registered case is making the association close to linear

```{r, echo = FALSE, warning = FALSE, message = FALSE}

ggplot( df , aes(x = confirmed, y = death ))  +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "Number of registered case (31 October 2020, ln scale)",y = "Number of registered death (ln scale)") +
  scale_x_continuous( trans = log_trans() )+
  scale_y_continuous( trans = log_trans() )
```

**level y - log x:**

Substantive reasoning: Level changes is harder to interpret, I would like to see how many people died per percentage change in confirmed case
Statistical reasoning: log transformation shows the better approximation

**log y - level x:**

Substantive reasoning: it does not give better interpretation 
Statistical reasoning: log transformation of number of registered death and taking level of number of case does not give good approximation


